### 2022-06-15 draft/draft.qmd
latest version

---
title: "Draft"
subtitle: "STA 210 - Project"
author: "Bayes' Harem - Christina Wang, Kat Cottrell, David Goh, Ethan Song"
format: pdf
editor: source
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
#| label: load-pkgs
#| message: false
#| warning: false
library(tidyverse)
library(tidymodels)
library(knitr)
library(ggfortify)
library(GGally)
```

```{r load-data}
#| message: false
abortion_data_full <- read_csv(here::here("data/abortion-attitudes", "wvs-usa-abortion-attitudes-data.csv"))
```

## Introduction and Research Question

# Motivation

Understanding public attitudes on divisive political issues is an important way for political leaders to mobilize voters and for lawmakers to draft laws that represent their constituents.  While it is simple to poll constituents’ positions on an issue, it can be challenging to assess the complex factors that influence and predict those stances.

Abortion is one such divisive issue in the United States, with strong organization of pro-choice and pro-life groups in states across the country in support of legislation for their respective sides (Ziegler, 2020). However, following the Supreme Court’s 1973 decision in Roe v. Wade, a ruling that has historically protected women's rights to abortion without excessive government restriction in America, the issue has risen in political salience. Both the pro-choice and pro-life movements have gained national prominence, and the two major political parties have polarized around the issue, with the Democratic Party in favor of and the Republican Party against policies legalization and increased access to abortion (Weinberger 2022). Abortion has also increasingly become a key issue that voters consider, with an increasing share of Americans identifying as “single-issue voters” regarding abortion (Brenan 2020).

In May 2022, a leaked draft opinion revealed that the US Supreme Court is prepared to overturn Roe v. Wade. Overturning Roe would dramatically change the trajectory of abortion politics in the US. Unless the US Congress passed a national policy, states would be able to decide whether or not to legalize abortion and gain much greater leverage in regulating access to the procedure (Weinberger 2022).

Given the potential overturning of Roe and the polarizing nature of the issue, it is important to understand how the American public feels about whether abortion should be legal or not, how accessible the procedure should be, and what factors influence these opinions. Understanding public opinion on the issue will ensure that political leaders are able to mobilize the correct constituencies, and that policy experts are able to pass policies on this issue that accurately reflect the preferences of the American people.

## Data description

We will conduct our analysis on data collected about Americans’ abortion attitudes, demographic information, and other ideologies, using an EDA-informed predictive model. The data are a nationally representative sample of the American people; hence we can infer that all the observations are independent and the variables may show a linear relationship. Given that this issue has become highly polarized by political party, we predict that liberal political attitudes and youth will correlate with belief that abortion is more justified.

The dataset observes "attitudes on the justifiability of abortion in the
United States across six waves of World Values Survey data" (README.md)
and some basic qualities of the respondents.

Observations include:

-   WVS country code

-   Generational wave (1982, 1990, 1995, 1999, 2006, or 2011)

-   Justifiability of abortion (1-10)

-   Age (17 to 96)

-   College graduate (1 for yes)

-   Female (1 for women) - Unemployed (1 = currently unemployed)

-   Ideology (1-10 for left-right)

-   Financial satisfaction (1-10 for least-most)

-   WVS post-materialist index (-1 = materialist. 2 = mixed. 3 =
    post-materialist)

-   Child autonomy index (-2 to 2 for obedience and religious
    faith-determination and independence)

-   Trust (1 = believes most people can be trusted)

-   Importance of God (1-10)

-   Opinion of respect for authority (-1-1 for bad to good)

-   National pride (1 = very proud to be an American)

The data were collected as part of the World Values Survey, which is
administered every few years and collects information about people's
values and beliefs worldwide. The survey aims to get a nationally
representative sample of a minimum of 1200 for most countries, and the
data are collected via face-to-face interviews at the respondents'
homes. The data included in this set specifically include responses from
6 waves of the survey (administered over the period 1982-2011). The
responses included in this set are from people in the United States, and
it examines their attitudes towards abortion.

## Research Question and Response Variable

In this project, we investigate the correlation between attitudes on the justifiability of abortion with demographic factors (such as age, gender, and education level) and personal attitudes towards other issues (such as political ideology, importance of religion, and respect for authority) in a representative sample of American citizens from 1982-2011.

The response variable, `Justifiability of abortion`, is a numerical measure on a scale of 1 to 10 on the individual person’s attitude toward whether abortion is justifiable or not.  Individuals responded 1 for “abortion is never justified” and 10 for “abortion is always justified.

## Exploratory Data Analysis

We create a visualization and summary statistics for the response variable.

```{r visualization-summary}
#| warning: false
ggplot(abortion_data, aes(x = aj)) +
  geom_histogram(binwidth = 1, aes(y=..density..)) + 
  geom_density(color = "red") +
  labs(title = "Histogram of abortion attitudes")

summary(abortion_data$aj)
```

We can observe from this histogram that the distribution of the outcome
variable is not a bell shape, and it is trimodal. This is likely because
the question's phrasing is similar to a yes/no question, but respondents
were asked to give their level of agreement on a scale of 1-10. This may
result in our model not being a good fit for the data if we attempt a
multiple linear regression model. We have two backup plans for this, if
our MLR eventually has a poor performance. First, we can truncate this
data into a categorical outcome variable such as (Agree, Disagree,
Undecided), and conduct a binomial or multinomial logistic regression.
Second, we can filter our population based on various characteristics
(if we have good reason to do so) so that our outcome data follows a
bell shape.

We now extend our exploratory data analysis (EDA) to some predictor
variables of interest; namely, the year, ideology, Child Autonomy Index,
Importance of God, Respect for Authority and National Pride predictors.
The EDA for each variable comprises a histogram and a boxplot of the
response variable grouped by predictor value. Additionally, we have a
jitter plot to explore the potential for an interaction effect between
year and importance of God.

```{r eda-year}
#| warning: false
#Year of survey
ggplot(abortion_data, aes(x = year)) +
  geom_histogram() + 
  labs(title = "Histogram of year")
ggplot(abortion_data, aes(group = year, y = aj)) +
  geom_boxplot() 
```

```{r eda-ideology}
#| warning: false
# ideology
ggplot(abortion_data, aes(x = ideology)) +
  geom_histogram() + 
  labs(title = "Histogram of ideology")
ggplot(abortion_data, aes(x = ideology,group = ideology, y = aj)) +
  geom_boxplot()
```

```{r eda-cai}
#| warning: false
# Child autonomy index
ggplot(abortion_data, aes(x = cai)) +
  geom_histogram() + 
  labs(title = "Histogram of Child autonomy index")
ggplot(abortion_data, aes(x = cai, group = cai, y = aj)) +
  geom_boxplot()
```

```{r eda-god}
#| warning: false
# Importance of God
ggplot(abortion_data, aes(x = godimportant)) +
  geom_histogram() + 
  labs(title = "Histogram of how respondent saw God's importance")
ggplot(abortion_data, aes(x = godimportant, group = godimportant, y = aj)) +
  geom_boxplot()
```

```{r eda-respect}
#| warning: false
# Respect for authority
ggplot(abortion_data, aes(x = respectauthority)) +
  geom_histogram() + 
  labs(title = "Histogram of respect for authority")
ggplot(abortion_data, aes(x = respectauthority, group = respectauthority, y = aj)) +
  geom_boxplot()
```

```{r eda-pride}
#| warning: false
# National Pride
ggplot(abortion_data, aes(x = nationalpride)) +
  geom_histogram() + 
  labs(title = "Histogram of National pride")
ggplot(abortion_data, aes(x = nationalpride, group = nationalpride, y = aj)) +
  geom_boxplot()
```

We now test if any of the predictors are strongly correlated with each
other.

```{r correlations}
#| warning: false
ggpairs(abortion_data, 
  columns = c("year", "ideology", "cai", "godimportant", "respectauthority", 
              "nationalpride"),
  columnLabels = c("year", "ideology", "child autonomy", "imptnce of God", "resp for autho", 
              "natl pride")) +
   theme(
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(angle = 45, size = 10),
    strip.text.y = element_text(angle = 0, hjust = 0)
    )
```

From these correlation matrices, we can conclude that the highest
correlations (above 0.2) are those between `godimportant` and `cai`
0.370 `godimportant` and `respectauthority` 0.244 `godimportant` and
`ideology` 0.243 `godimportant` and `nationalpride` 0.216
`nationalpride` and `respectauthority` 0.206 `respectauthority` and
`cai` 0.205

\pagebreak

## METHOD

We will conduct a Multiple Linear Regression (MLR) on the attitudes on
the justifiability of abortion against several other predictor variables
found in this data set. This is because our outcome variable, the
attitude towards abortion, is measured on a numeric scale from 1 to 10,
and there are multiple predictor variables that we feel are potentially
correlated with it.

We will now discuss the data cleaning and predictor selection process.
For observations, we removed the observations that were part of
"generational waves" before 1995, because many survey questions were
only added from 1995 onward, and observations before 1995 have a high
number of null values.

We removed the variable "wvsccode" as it is the same for all the
observations. The WVSC Code depends on the country the surveys were
conducted in, and all surveys in this data set were conducted in the
United States of America.

```{r filter-data}
abortion_data <- abortion_data_full %>% 
  filter(year >= "1995") %>%
  select(-starts_with("wave"), -starts_with("wvsccode"))
```

We chose to consider predictor variables that, from our exploratory data
analysis (EDA), looked like they were potentially correlated with the
outcome. Because the predictor variables were discrete, our EDA was
presented as a series of box plots. We observed the differences in
median and quartile values of abortion attitude across different
predictor values and ruled out variables with no visible hint of
relationship.

This led to our first set of predictor variables, which we designate as
Recipe1: 
- Year 
- Ideology 
- Child Autonomy Index 
- Importance of God 
- Respect for Authority 
- National Pride

We also wanted to choose predictor variables that are not strongly
correlated with each other, to avoid multicollinearity. From our EDA, we
found that several predictors had correlations of 0.2 or higher with
each other. Given these correlations, we decided to make a second recipe
(`Recipe2`) with the following predictor variables, after excluding
those with a high correlation with the `godimportant` variable:

-   Year
-   Importance of God

Finally, we considered that the distribution for the `godimportant`
variable is heavily left-skewed in our data set, which we observed in
the EDA. This may reduce the ability of a model that relies on the
"importance of God" predictor to explain the variations we observe in
the outcome. The same can be said of the `respectauthority` variable,
with more than 8 times the observations indicating "1" compared to the
observations indicating "-1".

Consequently, our third recipe (Recipe3) excludes `godimportant` and
uses the other variables that are not correlated with each other more
than 0.200:

-   Year
-   Ideology
-   Child Autonomy Index
-   National Pride

In the next section, we create and apply our recipes to fit MLR models.

## Data Split

## Fit Model

Note: make sure to create recipes, specify engine, make workflows, and
fit models here

## Inference

\pagebreak

##RESULTS

## Model Evaluation

## Model Comparison

## Model Diagnosis

## Discussion of Influential Points

## Data dictionary

The data dictionary can be found [here](ADD%20LINK!!!).

## References

Brenan, M. (2020, July 7). One in Four Americans Consider Abortion a Key
Voting Issue. Gallup. https://news.gallup.com/poll/313316/one-four-americans-consider-abortion-key-voting-issue.aspx

Weinberger, J. (2022, May 6). How we got here: Roe v. Wade from 1973 to
today. Vox. https://www.vox.com/23055389/roe-v-wade-timeline-abortion-overturn-political-polarization

Ziegler, M. (2020, October 22). Abortion politics polarized before Roe.
When it's gone, the fighting won't stop. The Washington Post. https://www.washingtonpost.com/outlook/2020/10/22/roe-polarize-abortion-politics/